Synergistic retrievals of ice in high clouds from elastic backscatter lidar, Ku-band radar and submillimeter wave radiometer observations

### 2023-08-22 RST/Submission/responseReviewers_Rev2.tex

\documentclass[12pt]{article}
\usepackage{amsmath,amssymb,amsthm}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{longtable}
\usepackage[margin=1in]{geometry}
\author{Mircea Grecu and John E. Yorks}
\title{Synergistic retrievals of ice in high clouds\\
Response to Reviewers\\Second Revision}
\date{}
\begin{document}
\maketitle

\noindent \textbf{Reviewer 2}\\
\noindent \textbf{General Summary}

@@

\noindent\textit{ 
For the sake of reproducibility, the authors should make all relevant source code publicly available. Furthermore, the data availability statement should clearly state which CloudSat products were used in the study.}\\

\noindent\textit{l. 23 - 24: Considering that the study neglects substantial uncertainties that will affect the accuracy of an actual synergistic retrieval, the sentence should mention that the 'significant improvements' were found under idealized conditions.}\\

\noindent Yes, we agree with the reviewer that the synergy was found under idealized conditions. We added a statement to this effect in the manuscript. Nevertheless, although additional work is likely needed to make optimal use of this synergy in practice, it is hard to envision how the tree instruments considered together cannot provide a more complete and quantitatively more accurate picture than any one of them considered independently.\\

@@

\noindent\textit{l. 85: You state that you use the 2C-ICE product to extend your CS-based estimates. However, you do not explain how.} \\

\noindent We apologize for the omission. When the CS-reflectivity is below the noise level, we use the IWC estimates in the 2C\-ICE product in conjunction with an $N_w$ provided by our statistical model. We added a sentence to the manuscript to clarify this point.\\

@@

\noindent\textit{l. 113: Please introduce the variable H as relative height above freezing level.}\\

Thank you for the suggestion. We introduced $H$ as the relative height above freezing level as suggested.\\

@@

\noindent\textit{l. 147: Here, you assume a fixed value for $N_w$. How is this consistent with the $N_w$ parametrization described earlier?}\\

\noindent We apologize for the confusion. As explained above, $IWC$ and $Z$ are normalized by $N_w$ in the derivation of $a_{norm}$. Consequently, the actual value of $N_w$ does not matter. One may set $N_w$ to the unit value. However, that would result in very unusual $IWC$ and $Z$, which would make sanity checks impossible.  We therefore set $N_w$ to a frequently used, but arbitrary values. This however does not have any impact on the actual value of $N_w$.\\

\noindent\textit{l. 258: How are absorption and extinction handled in your retrieval of IWC and PSD parameters from CS reflectivities?}\\

\noindent The gaseous absorption is modeled using the Rosenkranz (1998) model assuming the same water vapor, temperature and pressure profiles used in the lidar model and the radiative transfer calculations. The extinction is computed using the SSRGA theory, while the attenuation correction is handled through a forward attenuation correction procedure.
\\
\noindent\textit{l. 276 - 278: While generally true, this does not apply to this study since the test data stems from the same distribution as the a priori in either case.}\\

\noindent We just explain our preference for an observation-based dataset as opposed to a purely synthetic (CRM-based) dataset. However, we do not imply that our results are bias-free or that CRM can not eventually produce more complete and less-biased datasets. But until such datasets emerge, we consider the framework described.\\

@@

\noindent\textit{l. 291 - 292: Are you using a two-layer neural network without activations functions, or is this a typo? In the former case, I suggest you use a deeper neural network with non-linearities for the classification as you model may be underfitting the data.}\\

\noindent We apologize for the confusion and omissions. The neural network has two hidden layers with ReLU activations. We considered additional layers and/or different numbers of neurons per layer, but the results did not improve. We added a sentence to the manuscript to clarify this point. \\

\noindent\textit{l. 347: The new distribution of retrieved gravity centers cannot be described as narrow.}\\

\noindent We agree with the reviewer that the range of the retrieved gravity centers is not narrow. However, a significant fraction of retrieved IWC GC are within narrower range spanning from about 2.5 km above the freezing level to about 5.5 km above the freezing level. We added a sentence to the manuscript to clarify this point.\\

@@

\noindent\textit{Section 3. b.: Please add results for the lidar-radar retrievals.}\\

\noindent Thank you for the suggestion. We added the results for the lidar-radar retrievals in the revised manuscript.\\

@@

\noindent\textit{Table 2: Please include NRMS also for IWP and gravity centers. Since you mention it in the text, I suggest to also include the correlation coefficient in the table.}\\

\noindent\textit{l. 440: The term cross-validation implies several iterations of evaluations. This is different from what you are doing here.}\\

\noindent We agree with the reviewer that our particular form of validation, which is more generally known as hold-out validation, is not considered cross-validation by some authors. Nevertheless, other authors (e.g. Arlot 2010) consider hold-out the most basic form of cross-validation. To avoid controversy, 
we use term hold-out validation instead of cross-validation in the revised manuscript.\\

### 2023-08-22 RST/Submission/RSTManuscript_rev2.tex
revision 2

\abstract{ In this study, we investigate the synergy of elastic backscatter lidar, Ku-band radar, and sub-millimeter-wave radiometer measurements
in the retrieval of ice from satellite observations. The synergy is analyzed through the generation of a large dataset 
of Ice Water Content (IWC) profiles and simulated lidar, radar and radiometer observations. 
The characteristics of the instruments e.g. frequencies, sensitivities, etc. are set based on the expected characteristics 
of instruments of the Atmosphere Observing System (AOS) mission. A cross-validation methodology is used to assess the
accuracy of the IWC profiles retrieved from various combinations of observations from the three instruments. Specifically, the IWC and
associated observations are randomly divided into two datasets, one for training and the other for evaluation.
The training dataset is used to train the retrieval algorithm, while the evaluation dataset is used to assess the 
retrieval performance. The dataset of IWC profiles is derived from CloudSat 
reflectivity and CALIOP lidar observations.
The retrieval of the ice water content IWC profiles from the computed observations is achieved in two steps. 
In the first step, a class, out of 18 potential classes characterized by different vertical distribution of IWC, 
is estimated from the observations. The 18 classes are predetermined based on the k-Means clustering algorithm. 
In the second step, the IWC profile is estimated using an Ensemble Kalman Smoother (EKS) algorithm that uses the 
estimated class as a priori information.\\
The results of the study show that the synergy of lidar, radar, and radiometer observations is significant in the retrieval of the IWC profiles.  Nevertheless, it should be mentioned that this synergy was found under idealized conditions, and additional work might be required to materialize it in practice. The inclusion of the lidar backscatter observations in the retrieval process has a larger impact on the retrieval performance than the inclusion of the radar observations. As ice clouds have a significant impact on atmospheric radiative processes, this work is relevant to ongoing efforts to reduce uncertainties in climate analyses and projections.}

@@

The future NASA Atmospheric Observing System (AOS) mission \citep{braun2022} is expected to feature new combinations of observations that may be used to quantify the amounts of ice in high clouds and characterize the microphysical properties of ice particles. In the AOS terminology, high clouds are convectively generated clouds \citep{braun2022}. They are the result of strong vertical mass transport accompanied by
horizontal transport of hydrometeors in the upper troposphere.  As high clouds are of paramount importance in understanding the quantifying radiative processes, they constitute one of the AOS major objectives \citep{braun2022}. To achieve its objectives, AOS will rely on a combination of active and passive observations onboard multiple spacecrafts in two different orbits.  One of the orbits is expected to be polar, while the other is expected to be inclined, which would allow the study of atmospheric processes at sub-daily time scales, with an emphasis on deep convection, high clouds, and aerosols \citep{braun2022,yorks2022}. The AOS inclined observations will include backscatter from an elastic backscatter lidar \citep{weitkamp2006}, Ku-band radar reflectivity, and submillimeter wave radiometer brightness temperature measurements. 
While not necessarily optimal for cloud ice estimation, these measurements are complimentary and enable the synergistic characterization of ice clouds. That is, despite the fact that lidar observations attenuate quickly in thick ice clouds and the Ku-band radar will not be able to detect clouds 
characterized by an echo weaker than 8.0 dBZ (which is the expected sensitivity of the radar in the inclined
orbit), the active observations are expected to provide context 
that may be incorporated into the radiometer retrievals. Herein, the term retrieval is defined as the process of estimating geophysical variables from remote sensing observations.

@@

In this study, we investigate the impact of 
incorporating lidar and radar observations into the radiometer retrieval of ice clouds. Because the existing amount of coincident backscatter lidar, Ku-band radar, and submillimeter-wave  radiometer observations is rather insufficient to derive conclusive results, we employ accurate 
physical models to simulate lidar, radar and radiometer observations and use a cross-validation methodology to characterize the retrieval accuracy. As estimates from passive instrument observations strongly depend on "a priori" information \citep{rodgers2000inverse}, for results to be 
relevant in real applications, it is necessary to base them on realistic vertical distributions of ice properties.  Such distributions may be derived from cloud-resolving-model (CRM) simulations \citep{pfreundschuh2020synergistic,liu2022assessing} or directly from observations.  In this study, we employ the latter approach, as CRMs may still be deficient in properly reproducing the vertical distribution of ice clouds and their associated microphysical properties.  Specifically, we use observations and products from the CloudSat (CS) mission (Stephens et al. 2002) to derive a database of ice microphysical properties and associated simulated lidar, radar, and radiometer observations.  To account for variability in the Particle Size Distributions (PSD) that may not be well represented in the 2C-ICE product, we use a simple but effective approach to perturb the PSD generalized intercepts from their nominal values.
The resulting database is used to investigate the accuracy of estimated ice cloud properties from the simulated observations. Another major difference relative to previous studies is the unique combination of instruments investigated herein. It should be mentioned that although based on observations rather than CRM simulations, the approach used in this study is
idealized in many respects. As a consequence, the results presented may not unbiasedly extrapolate to practice.
Nonetheless, the results are expected to provide useful insights into the potential of the AOS inclined observations and
good first step towards the development of an operational synergistic retrieval of ice clouds from AOS inclined observations. The article is organized as follows.  In Section 2, we describe the approach used to derive the ice properties and the associated simulated observations, the retrieval and the evaluation methodology.  In Section 3, we present the results of the evaluation methodology. We conclude in Section 4.

% \subsection*{subsection}
% text...
\section{Methodology}
As previously mentioned, we use CloudSat (CS) observations \citep{stephens2002} to derive the vertical 
distributions of ice properties needed in the investigation.  Although research quality CS cloud ice 
products exist, to maximize the physical consistency of the approach, we do not use them but derive ice amounts
and associated properties directly from CS reflectivity observations.  This ensures the consistency between 
the particle
distribution assumptions and the electromagnetic scattering properties used in the CS reflectivity processing 
and those used the simulation of the lidar, Ku-band radar and radiometer observations.  
Our CS-based ice product is optimized to be consistent with the synergistic 
Cloudsat and CALIPSO Ice Cloud Property Product (2C-ICE) of \cite{deng2015}.  When the CALIPSO lidar detects echo
associated with clouds but the CS radar signal is below the noise level, we use the 2C-ICE product to extend our 
CS-based estimates. Lidar, Ku-band radar, and submillimeter-wave radiometer observations are simulated 
from CS observations using accurate physical 
models and realistic assumptions consistent with the most recent knowledge in the field of ice cloud 
microphysics, and a non-parametric estimation methodology based on the k-Means clustering algorithm 
\cite{mackay2003information} is used to 
investigate the instrument synergy.  Details of the methodology are presented below.

\subsection{Assumptions and forward models}
To quantify the number of ice particles in an elementary atmospheric volume as a function of their size, we use
normalized gamma functions \citep{bringi2003}.  The benefit of normalized gamma functions is that they 
encapsulate the variability of Ice Water Content (IWC) - reflectivity relationship into a single parameter,
i.e. the 
normalized Particle Size Distribution (PSD) intercept \citep{testud2001,bringi2003}. The normalized
PSD intercept is defined as $N_w=\frac {4^4} {\pi \rho_w} \frac {IWC} {D_m^4}$, where $IWC$ is the ice water 
content associated with the PSD, and $D_m$ is the mass weighted mean diameter.  Based on the work of \cite{testud2001},
\cite{ferreira2001} and \cite{delanoe2014} showed 
that the variability in IWC reflectivity (Z) relationships may be fully explained by variability in $N_w$, and 
that a formula of the type
\begin{equation}
IWC=N_w^{1-b}a_{norm} Z^b \label{eq:1}
\end{equation}
(where $a_{norm}$ and $b$ are constants) explains almost perfectly the relationships between IWC and Z calculated from observed PSDs.
However, equation (\ref{eq:1}) is not 
sufficient to derive accurate, unbiased estimates of ice water contents, because $N_w$ varies considerably in 
time and space. Nevertheless, multiple studies showed that it is beneficial to parameterize $N_w$ as a function 
of various variables, such as temperature \citep{hogan2006retrieval, delanoe2008variational,deng2010tropical}, 
rather than using $N_w$ independent reflectivity ice relations.  In this study, we parameterize
$N_w$ as a function of temperature based on the CloudSat 2C-ICE product \citep{deng2010tropical,deng2013evaluation}.  
A scatter-plot analysis of relationships between the 2C-ICE IWC and the associated reflectivity suggests that
the multiplicative coefficient in a power low ice-reflectivity relationship is parameterized as a function of temperature 
(or equivalently, as a function of height relative to the freezing level) in 
the default 2C-ICE retrievals. The default multiplicative coefficient (the value the provides the IWC estimate prior
to the ingestion of the lidar observations) may be simply estimated by regressing the 2C-ICE IWC on the associated CloudSat reflectivity 
as a function of height relative the freezing level. Specifically, a regression of the type $ln(IWC)=ln(a)+b*ln(Z)$ is performed for 
each bin referenced relative to the zero degree bin is performed, and the variation of $ln(a)$ is parameterized as a function of the
relative height.  The results are shown in Figure \ref{f2}. As apparent in the figure, and as expected, $a$ exhibits a strong variation 
with the relative height. Given that any deviation of the multiplicative coefficient in 
an IWC-Z relation from an average is equivalent to a deviation of the associated $N_w$ from its mean value 
\citep{ferreira2001,delanoe2014}, the variation of $a$ as a function of relative-height may be converted into an $N_w$ 
as a function of relative-height relationship.  We, therefore, use the data in Figure \ref{f2} to parameterize $N_w$ as a 
function of the relative height.  Specifically, given that $(1-b)ln(N_w)+ln(a_{norm})=ln(a)$ and $ln(a)=ln(a_0)+s*H$, with 
$a_0$ and $s$ the intercept and the slope of a regression for the data in Figure \ref{f2}, we 
can express $N_w$ as a function of the relative height, slope $s$, and parameters $a_0$ and $a_{norm}$.

@@

Moreover, the SSRGA theory was developed for millimeter and 
submillimeter-wave calculations and may not be applicable at lidar's wavelength. 
Therefore, lidar observations are computed using the Mie solution included in the scatter-1.1 
package for a backscatter lidar, given that such a lidar is expected to be flown with a 
Ku-band radar and microwave radiometer as part of the NASA Earth System Observatory 
(ESO) Atmosphere Observing System (AOS) in an inclined and/or polar orbit \citep{yorks2022}.
While uncertainties in the lidar forward model are rather complex and difficult to quantify,
uncertainties in the lidar observations may be accounted for to some extent by including a
multiplicative random noise factor in the calculated lidar backscatter values.  
Specifically, the calculated lidar backscatter values are multiplied with a log-normally distributed random variable 
with 0.0 mean and a standard deviation of 0.1. This results in uncertainties of about 10\% in the lidar backscatter values, 
conservative compared to the expected performance of the AOS lidars at 532 nm but consistent the performance of 
Cloud-Aerosol Transport System (CATS) lidar system at 1064 nm reported in \citep{pauly2019}. While the AOS lidars 
will have expected backscatter accuracies of 2-5\%, the lidar simulations in this paper are rather idealized in that 
they do not include random errors due to daytime solar background noise or systematic errors such as calibration accuracy 
or lidar ratio assumptions. More advanced models of the observation errors exist \citep{liu2006}, but are not 
considered here. There are three main reasons this approach was taken: (1) it is important to
understand the limitations of combining these three datasets before factoring in 
individual sensor limitations, (2) while more accurate calculations based on more 
realistic ice particle shapes exist, they are rather incomplete and not readily available,
and (3) Wagner and Delene (2022) compared lidar backscatter observations with backscatter 
calculations based on coincident PSD observations and the Mie solution and found good 
agreement, which suggests that electromagnetic properties derived from Mie calculations 
are adequate for practical applications. 
The lidar molecular backscatter and extinction
are calculated using the lidar module of the CFMIP Observation Simulator Package (COSP; 
Bodas-Salcedo et al. (2011)). COSP simulates the lidar total attenuated backscatter 
signal and scattering ratios at 532 nm for scenes with and without clouds. COSP also 
assumes cloud particles are spherical so that the backscattering phase function is 
estimated based on the effective radius (Mie theory). Despite this assumption, the 
COSP simulations agree well with CALIOP observations. To account for multiple-scattering 
in the lidar observations, we are using the multiscatter-1.2.11 model \citep{multiscatter-1.2.11} of 
\cite{hogan2008lidar_ms}.

Shown in Figure 3 are the distributions of simulated Ku-band radar reflectivity and 
lidar backscatter as function of height above the freezing level.
As apparent in the figure, the IWCs associated with detectable Ku-band reflectivity
signal are likely to
occur near mostly around 3.0 km above the freezing level, while the lidar backscatter
distribution exhibits a peak at around 6.0-7.0 km
above the freezing level, which is consistent with the fact the lidar observations
are strongly attenuated in the bottom part of the cloud.  It should be noted that, consistently with the objective of
 this study, only CS reflectivity profiles with no echo at or below the freezing level were selected
and used in the calculation of Ku-band reflectivity distributions shown in Figure \ref{f3}.
The vertical resolutions of
the radar and lidar observations are the same, i.e. 240 m, and the same as the resolution of the
CS observations upon which they are based. In reality, the radar and lidar observations are likely to have
different vertical resolutions and footprint sizes, which would need to be accounted for in a more 
realistic study.  It should be noted though that the AOS radars and radiometers are expected to achieve
Nyquist sampling, which enables resolution enhancement \citep{early2001image}.  Nevertheless, differences
in the vertical resolutions of the radar and lidar observations and differences in the footprint sizes
may deteriorate the performance of the retrieval algorithm.

@@

The processing steps used to process the CS reflectivity observations and calculate the lidar, Ku-band and 
submillimeter-wave radiometer observations may be summarized as follows:
\begin{enumerate}
\item Derivation of physically consistent radar and radiometer lookup tables to relate basic radar and 
radiometer properties (e.g. reflectivity, attenuation, extinction, scattering-albedo, etc.) to PSD parameters 
such as $IWC$ and $D_m$. The tables are derived for a single of $N_w$, but are usable with any value of 
$N_w$ using the "normalization" operations described in \citep{grecu2011}.
\item Derivation of $N_w$-relative height parameterization using the 2C-ICE product.
\item Estimation of IWC and related PSD parameters from CS W-band radar observations, using the tables 
constructed Step 1, and $N_w$ profiles derived through parameterization obtained in Step 2 and perturbed using the random model 
described above. Specifically, given $N_w$ and the associated $Z$ at W-band, we estimate $IWC/N_w$ as a function of $Z/N_w$ using the
normalized tables and then derive $IWC$ from $IWC/N_w$ and $N_w$.
\item Calculation of lidar, Ku-band radar and radiometer observations from the estimates derived in Step 3 
and the tables obtained in Step 1, and extended with non-zero 2C-ICE estimates for the radar bins characterized by no echo
above the noise level.
\end{enumerate}

@@

% text...
% 
\section{Conclusions}
In this study, we investigate the synergy of elastic backscatter lidar, Ku-band radar, and sub-millimeter-wave radiometer
measurements in the retrieval of the ice from satellite observations.  The synergy is analyzed through
the generation of a large dataset of IWC profile and the calculation of lidar, radar and radiometer 
observations using realistic models. The characteristics of the instruments (e.g. frequencies, sensitivities,
etc.) are set based on the expected characteristics of instruments of the AOS mission. 
A cross-validation methodology is used to assess the 
accuracy of the retrieved IWC profiles from various combinations of observations from the three instruments.
Specifically, the IWC and associated observations is randomly divided into two datasets, one for
training and the other for evaluation.  The training dataset is used to train the retrieval algorithm,
while the evaluation dataset is used to assess the retrieval performance.

>>>>

\abstract{ In this study, we investigate the synergy of elastic backscatter lidar, Ku-band radar, and sub-millimeter-wave radiometer measurements in the retrieval of ice from satellite observations. The synergy is analyzed through the generation of a large dataset of Ice Water Content (IWC) profiles and simulated lidar, radar and radiometer observations. The characteristics of the instruments e.g. frequencies, sensitivities, etc. are set based on the expected characteristics of instruments of the Atmosphere Observing System (AOS) mission. A hold-out validation methodology is used to assess the accuracy of the IWC profiles retrieved from various combinations of observations from the three instruments. Specifically, the IWC and associated observations are randomly divided into two datasets, one for training and the other for evaluation. The training dataset is used to train the retrieval algorithm, while the evaluation dataset is used to assess the retrieval performance. The dataset of IWC profiles is derived from CloudSat reflectivity and CALIOP lidar observations. The retrieval of the ice water content IWC profiles from the computed observations is achieved in two steps. In the first step, a class, out of 18 potential classes characterized by different vertical distribution of IWC, is estimated from the observations. The 18 classes are predetermined based on the k-Means clustering algorithm. In the second step, the IWC profile is estimated using an Ensemble Kalman Smoother (EKS) algorithm that uses the estimated class as a priori information.\\
The results of the study show that the synergy of lidar, radar, and radiometer observations is significant in the retrieval of the IWC profiles.  Nevertheless, it should be mentioned that this synergy was found under idealized conditions, and additional work might be required to materialize it in practice. The inclusion of the lidar backscatter observations in the retrieval process has a larger impact on the retrieval performance than the inclusion of the radar observations. As ice clouds have a significant impact on atmospheric radiative processes, this work is relevant to ongoing efforts to reduce uncertainties in climate analyses and projections.}

@@

The future NASA Atmospheric Observing System (AOS) mission \citep{braun2022} is expected to feature new combinations of observations that may be used to quantify the amounts of ice in high clouds and characterize the microphysical properties of ice particles. In the AOS terminology, high clouds are convectively generated clouds \citep{braun2022}. They are the result of strong vertical mass transport accompanied by horizontal transport of hydrometeors in the upper troposphere.  As high clouds are of paramount importance in understanding the quantifying radiative processes, they constitute one of the AOS major objectives \citep{braun2022}. To achieve its objectives, AOS will rely on a combination of active and passive observations onboard multiple spacecrafts in two different orbits.  One of the orbits is expected to be polar, while the other is expected to be inclined, which would allow the study of atmospheric processes at sub-daily time scales, with an emphasis on deep convection, high clouds, and aerosols \citep{braun2022,yorks2022}. The AOS inclined observations will include backscatter from an elastic backscatter lidar \citep{weitkamp2006}, Ku-band radar reflectivity, and submillimeter wave radiometer brightness temperature measurements. 
While not necessarily optimal for cloud ice estimation, these measurements are complimentary and enable the synergistic characterization of ice clouds. That is, despite the fact that lidar observations attenuate quickly in thick ice clouds and the Ku-band radar will not be able to detect clouds characterized by an echo weaker than 8.0 dBZ (which is the expected sensitivity of the radar in the inclined orbit), the active observations are expected to provide context that may be incorporated into the radiometer retrievals. Herein, the term retrieval is defined as the process of estimating geophysical variables from remote sensing observations.

@@

In this study, we investigate the impact of incorporating lidar and radar observations into the radiometer retrieval of ice clouds. Because the existing amount of coincident backscatter lidar, Ku-band radar, and submillimeter-wave  radiometer observations is rather insufficient to derive conclusive results, we employ accurate physical models to simulate lidar, radar and radiometer observations and use a hold-out validation methodology to characterize the retrieval accuracy. As estimates from passive instrument observations strongly depend on "a priori" information \citep{rodgers2000inverse}, for results to be relevant in real applications, it is necessary to base them on realistic vertical distributions of ice properties.  Such distributions may be derived from cloud-resolving-model (CRM) simulations \citep{pfreundschuh2020synergistic,liu2022assessing} or directly from observations.  In this study, we employ the latter approach, as CRMs may still be deficient in properly reproducing the vertical distribution of ice clouds and their associated microphysical properties.  Specifically, we use observations and products from the CloudSat (CS) mission (Stephens et al. 2002) to derive a database of ice microphysical properties and associated simulated lidar, radar, and radiometer observations.  To account for variability in the Particle Size Distributions (PSD) that may not be well represented in the 2C-ICE product, we use a simple but effective approach to perturb the PSD generalized intercepts from their nominal values. The resulting database is used to investigate the accuracy of estimated ice cloud properties from the simulated observations. Another major difference relative to previous studies is the unique combination of instruments investigated herein. It should be mentioned that although based on observations rather than CRM simulations, the approach used in this study is idealized in many respects. As a consequence, the results presented may not unbiasedly extrapolate to practice. Nonetheless, the results are expected to provide useful insights into the potential of the AOS inclined observations and good first step towards the development of an operational synergistic retrieval of ice clouds from AOS inclined observations. The article is organized as follows.  In Section 2, we describe the approach used to derive the ice properties and the associated simulated observations, the retrieval and the evaluation methodology.  In Section 3, we present the results of the evaluation methodology. We conclude in Section 4.

% \subsection*{subsection}
% text...
\section{Methodology}
As previously mentioned, we use CloudSat (CS) observations \citep{stephens2002} to derive the vertical distributions of ice properties needed in the investigation.  Although research quality CS cloud ice products exist, to maximize the physical consistency of the approach, we do not use them but derive ice amounts and associated properties directly from CS reflectivity observations.  This ensures the consistency between the particle distribution assumptions and the electromagnetic scattering properties used in the CS reflectivity processing and those used the simulation of the lidar, Ku-band radar and radiometer observations. Our CS-based ice product is optimized to be consistent with the synergistic Cloudsat and CALIPSO Ice Cloud Property Product (2C-ICE) of \cite{deng2015}.  When the CALIPSO lidar detects echo associated with clouds but the CS radar signal is below the noise level, we use the 2C-ICE product to extend our CS-based estimates. Specifically, we use the IWC estimates from the 2C-ICE product in conjunction with a Particle Size Distribution (PSD) generalized intercept \citep{testud2001} provided by statistical model described in the next section to characterize the ice PSDs of clouds with radar echo below the noise level of the CloudSat radar. Lidar, Ku-band radar, and submillimeter-wave radiometer observations are simulated from CS observations using accurate physical models and realistic assumptions consistent with the most recent knowledge in the field of ice cloud microphysics, and a non-parametric estimation methodology based on the k-Means clustering algorithm \cite{mackay2003information} is used to investigate the instrument synergy.  Details of the methodology are presented below.

\subsection{Assumptions and forward models}
To quantify the number of ice particles in an elementary atmospheric volume as a function of their size, we use normalized gamma functions \citep{bringi2003}.  The benefit of normalized gamma functions is that they encapsulate the variability of Ice Water Content (IWC) - reflectivity relationship into a single parameter, i.e. the normalized Particle Size Distribution (PSD) intercept \citep{testud2001,bringi2003}. The normalized PSD intercept is defined as $N_w=\frac {4^4} {\pi \rho_w} \frac {IWC} {D_m^4}$, where $IWC$ is the ice water content associated with the PSD, and $D_m$ is the mass weighted mean diameter.  Based on the work of \cite{testud2001}, \cite{ferreira2001} and \cite{delanoe2014} showed that the variability in IWC reflectivity (Z) relationships may be fully explained by variability in $N_w$, and that a formula of the type
\begin{equation}
IWC=N_w^{1-b}a_{norm} Z^b \label{eq:1}
\end{equation}
(where $a_{norm}$ and $b$ are constants) explains almost perfectly the relationships between IWC and Z calculated from observed PSDs.However, equation (\ref{eq:1}) is not  sufficient to derive accurate, unbiased estimates of ice water contents, because $N_w$ varies considerably in time and space. Nevertheless, multiple studies showed that it is beneficial to parameterize $N_w$ as a function  of various variables, such as temperature \citep{hogan2006retrieval, delanoe2008variational,deng2010tropical}, rather than using $N_w$ independent reflectivity ice relations.  In this study, we parameterize $N_w$ as a function of temperature based on the CloudSat 2C-ICE product \citep{deng2010tropical,deng2013evaluation}. A scatter-plot analysis of relationships between the 2C-ICE IWC and the associated reflectivity suggests that the multiplicative coefficient in a power low ice-reflectivity relationship is parameterized as a function of temperature (or equivalently, as a function of height relative to the freezing level) in the default 2C-ICE retrievals. The default multiplicative coefficient (the value the provides the IWC estimate prior to the ingestion of the lidar observations) may be simply estimated by regressing the 2C-ICE IWC on the associated CloudSat reflectivity as a function of $H$, defined as height relative the freezing level. Specifically, a regression of the type $ln(IWC)=ln(a)+b*ln(Z)$ is performed for each bin referenced relative to the zero degree bin is performed, and the variation of $ln(a)$ is parameterized as a function of the relative height.  The results are shown in Figure \ref{f2}. As apparent in the figure, and as expected, $a$ exhibits a strong variation with the relative height. Given that any deviation of the multiplicative coefficient in an IWC-Z relation from an average is equivalent to a deviation of the associated $N_w$ from its mean value \citep{ferreira2001,delanoe2014}, the variation of $a$ as a function of relative-height may be converted into an $N_w$ as a function of relative-height relationship.  We, therefore, use the data in Figure \ref{f2} to parameterize $N_w$ as a function of the relative height.  Specifically, given that $(1-b)ln(N_w)+ln(a_{norm})=ln(a)$ and $ln(a)=ln(a_0)+s*H$, with $a_0$ and $s$ the intercept and the slope of a regression for the data in Figure \ref{f2}, we can express $N_w$ as a function of the relative height, slope $s$, and parameters $a_0$ and $a_{norm}$.

@@

Moreover, the SSRGA theory was developed for millimeter and submillimeter-wave calculations and may not be applicable at lidar's wavelength. Therefore, lidar observations are computed using the Mie solution included in the scatter-1.1 package for a backscatter lidar, given that such a lidar is expected to be flown with a Ku-band radar and microwave radiometer as part of the NASA Earth System Observatory (ESO) Atmosphere Observing System (AOS) in an inclined and/or polar orbit \citep{yorks2022}.While uncertainties in the lidar forward model are rather complex and difficult to quantify, uncertainties in the lidar observations may be accounted for to some extent by including a multiplicative random noise factor in the calculated lidar backscatter values. Specifically, the calculated lidar backscatter values are multiplied with a log-normally distributed random variable with 0.0 mean and a standard deviation of 0.1. This results in uncertainties of about 10\% in the lidar backscatter values, conservative compared to the expected performance of the AOS lidars at 532 nm but consistent the performance of Cloud-Aerosol Transport System (CATS) lidar system at 1064 nm reported in \citep{pauly2019}. While the AOS lidars will have expected backscatter accuracies of 2-5\%, the lidar simulations in this paper are rather idealized in that they do not include random errors due to daytime solar background noise or systematic errors such as calibration accuracy or lidar ratio assumptions. More advanced models of the observation errors exist \citep{liu2006}, but are not considered here. There are three main reasons this approach was taken: (1) it is important to understand the limitations of combining these three datasets before factoring in individual sensor limitations, (2) while more accurate calculations based on more realistic ice particle shapes exist, they are rather incomplete and not readily available, and (3) Wagner and Delene (2022) compared lidar backscatter observations with backscatter calculations based on coincident PSD observations and the Mie solution and found good agreement, which suggests that electromagnetic properties derived from Mie calculations are adequate for practical applications. The lidar molecular backscatter and extinction are calculated using the lidar module of the CFMIP Observation Simulator Package (COSP; Bodas-Salcedo et al. (2011)). COSP simulates the lidar total attenuated backscatter signal and scattering ratios at 532 nm for scenes with and without clouds. COSP also assumes cloud particles are spherical so that the backscattering phase function is estimated based on the effective radius (Mie theory). Despite this assumption, the COSP simulations agree well with CALIOP observations. To account for multiple-scattering in the lidar observations, we are using the multiscatter-1.2.11 model \citep{multiscatter-1.2.11} of \cite{hogan2008lidar_ms}.

Shown in Figure 3 are the distributions of simulated Ku-band radar reflectivity and lidar backscatter as function of height above the freezing level. As apparent in the figure, the IWCs associated with detectable Ku-band reflectivity signal are likely to occur near mostly around 3.0 km above the freezing level, while the lidar backscatter distribution exhibits a peak at around 6.0-7.0 km above the freezing level, which is consistent with the fact the lidar observations are strongly attenuated in the bottom part of the cloud.  It should be noted that, consistently with the objective of this study, only CS reflectivity profiles with no echo at or below the freezing level were selected and used in the calculation of Ku-band reflectivity distributions shown in Figure \ref{f3}. The vertical resolutions of the radar and lidar observations are the same, i.e. 240 m, and the same as the resolution of the CS observations upon which they are based. In reality, the radar and lidar observations are likely to have different vertical resolutions and footprint sizes, which would need to be accounted for in a more realistic study.  It should be noted though that the AOS radars and radiometers are expected to achieveNyquist sampling, which enables resolution enhancement \citep{early2001image}.  Nevertheless, differences in the vertical resolutions of the radar and lidar observations and differences in the footprint sizes may deteriorate the performance of the retrieval algorithm.

@@

The processing steps used to process the CS reflectivity observations and calculate the lidar, Ku-band and submillimeter-wave radiometer observations may be summarized as follows:
\begin{enumerate}
\item Derivation of physically consistent radar and radiometer lookup tables to relate basic radar and radiometer properties (e.g. reflectivity, attenuation, extinction, scattering-albedo, etc.) to PSD parameters such as $IWC$ and $D_m$. The tables are derived for a single of $N_w$, but are usable with any value of 
$N_w$ using the "normalization" operations described in \citep{grecu2011}.
\item Derivation of $N_w$-relative height parameterization using the 2C-ICE product.
\item Estimation of IWC and related PSD parameters from CS W-band radar observations, using the tables constructed Step 1, and $N_w$ profiles derived through parameterization obtained in Step 2 and perturbed using the random model described above. Specifically, given $N_w$ and the associated $Z$ at W-band, we estimate $IWC/N_w$ as a function of $Z/N_w$ using the normalized tables and then derive $IWC$ from $IWC/N_w$ and $N_w$.  A simple forward attenuation correction \citep{meneghini1978} is applied to the W-band reflectivity observations prior to the estimation of $IWC$.  Attenuation due to water vapor and other gases is calculated using the Rosenkranz model \citep{rosenkranz1998} assuming the same temperature, pressure and temperature profiles used in the radiometer simulations.

\item Calculation of lidar, Ku-band radar and radiometer observations from the estimates derived in Step 3 and the tables obtained in Step 1, and extended with non-zero 2C-ICE estimates for the radar bins characterized by no echo
above the noise level.
\end{enumerate}

@@

% text...
% 
\section{Conclusions}
In this study, we investigate the synergy of elastic backscatter lidar, Ku-band radar, and sub-millimeter-wave radiometer measurements in the retrieval of the ice from satellite observations.  The synergy is analyzed through the generation of a large dataset of IWC profile and the calculation of lidar, radar and radiometer observations using realistic models. The characteristics of the instruments (e.g. frequencies, sensitivities, etc.) are set based on the expected characteristics of instruments of the AOS mission. A hold-out validation methodology is used to assess the accuracy of the retrieved IWC profiles from various combinations of observations from the three instruments. Specifically, the IWC and associated observations is randomly divided into two datasets, one for training and the other for evaluation.  The training dataset is used to train the retrieval algorithm, while the evaluation dataset is used to assess the retrieval performance.
